---
title: "Estimation of EWAC weights"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Loading data
source("../01_src/functions.R")
source("../01_src/filepaths.R")
source("../01_src/audit_weights.R")

library(tidyverse)
library(rstan)

# library("knitr")
# library("xtable")
# library("stargazer")

load(file.path("..", "02_data", "ATS.rda"))
ats <- ats %>% 
  # filter(actage >= 18) %>%
  filter(audit1 > 1 & audit1 < 7) %>% 
  filter(!audit2_label %in% c("Refused", "Don't know")) %>% 
  filter(!audit3_label %in% c("Refused", "Don't know")) %>% 
  mutate(audit1_label = factor(audit1_label),
         audit2_label = factor(audit2_label), 
         audit3_label = factor(audit3_label))

ats$ewac_mp <- proc_EWAC(
  ats[,c("audit1_label", "audit2_label", "audit3_label")],
  audit_coef = audit_weights_midpoints, method = "qfv")

ats_subset <- ats %>% 
  transmute(GF = gfmeanweekly,
            audit1_ = as.factor(audit1),
            audit2_ = as.factor(audit2),
            audit3_ = as.factor(audit3)) %>% 
  model.frame(data = ., formula = GF ~ audit1_ + audit2_ + audit3_)

stan_input <- data.frame(
  model.matrix(~ audit1_ + audit2_ + audit3_, ats_subset, 
                   contrasts = list(
                     audit1_ = contrasts(ats_subset$audit1_, contrasts = F),
                     audit2_ = contrasts(ats_subset$audit2_, contrasts = F),
                     audit3_ = contrasts(ats_subset$audit3_, contrasts = F)
                   ))
)[,-1]

stan_input$GF <- ats_subset$GF
stan_input <- na.omit(stan_input)

stan_input <- as.list(stan_input)
stan_input$N <- length(stan_input$audit1_2)
rm(ats_subset)
```

The ATS dataset provides both GF and Extended AUDIT-C measures for a sample of household residents in England between waves 110 and 133, that is, November 2015 and October 2017. We restrict the analysis to adults aged 18 years and over. The GF schedule was not administered to participants responding 'Never' in AUDIT item 1. Consequently, this analysis is conducted only on drinkers.


```{r cars}
library(rstan)
mod1_fit <- stan(
  file = "stan_model.stan",  # Stan program
  data = stan_input,    # named list of data
  chains = 2,             # number of Markov chains
  warmup = 500,          # number of warmup iterations per chain
  iter = 1000,            # total number of iterations per chain
  cores = 2,              # number of cores (could use one per chain)
  init = list(
    list(sigma = 10, binge = 6, 
         F2 = 0.173, F3 = 0.69, F4 = 2.5, F5 = 4.5, F6 = 6.5, 
         Q1 = 1.5, Q2 = 3.5, Q3 = 5.5, Q4 = 8, Q5 = 11, Q6 = 14, Q7 = 20, 
         V1 = 0, V2 = 0.3, V3 = 0.7, V4 = 2, V5 = 5),
    list(sigma = 12, binge = 8, 
         F2 = 0.192, F3 = 0.767, F4 = 2.667, F5 = 4.667, F6 = 6.667, 
         Q1 = 1.667, Q2 = 3.667, Q3 = 5.667, Q4 = 8.333, Q5 = 11.333, Q6 = 14.333, Q7 = 21, 
         V1 = 0, V2 = 0.4, V3 = 0.8, V4 = 2.5, V5 = 6)
  ))

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
results <- summary(mod1_fit, pars = c('F2', 'F3', 'F4', 'F5', 'F6', 'Q1', 'Q2', 'Q3', 'Q4', 'Q5', 'Q6', 'Q7', 'V1', 'V2', 'V3', 'V4'))$summary

write.csv(results, file = "results1000its.csv")
results
```

```{r}
plot(mod1_fit, plotfun = "trace", pars = c('F2', 'F3', 'F4', 'F5', 'F6', 'Q1', 'Q2', 'Q3', 'Q4', 'Q5', 'Q6', 'Q7', 'V1', 'V2', 'V3', 'V4'))
```
 
 
 
```{r}

library(randomForest)
rf1 <- ats %>% 
  select(gfmeanweekly, audit1_label, audit2_label, audit3_label) %>% 
  na.omit() %>%
  randomForest(gfmeanweekly ~ audit1_label + audit2_label + audit3_label,
             data = ., ntree = 50)

ats$ewac_rf <- predict(rf1, newdata = ats)

save(rf1, file = "../02_data/random_forest_model.rda")

```
 
