---
title: "Estimation of EWAC weights"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```

The ATS dataset provides both GF and Extended AUDIT-C measures for a sample of household residents in England between waves 110 and 133, that is, November 2015 and October 2017. We restrict the analysis to adults aged 18 years and over. The GF schedule was not administered to participants responding 'Never' in AUDIT item 1. Consequently, this analysis is conducted only on drinkers.


```{r cars}

# library("knitr")
# library("xtable")
# library("stargazer")


```

## Including Plots


 
```{r}

library(randomForest)
rf1 <- ats %>% 
  select(gfmeanweekly, audit1_label, audit2_label, audit3_label) %>% 
  na.omit() %>%
  randomForest(gfmeanweekly ~ audit1_label + audit2_label + audit3_label,
             data = ., ntree = 50)

ats$ewac_rf <- predict(rf1, newdata = ats)

save(rf1, file = "../02_data/random_forest_model.rda")

```
## Priors


```{r}
# prior for binge intorduced on 25/09/19
plot(seq(0, 20, .01), dgamma(seq(0, 20, .01)-5, shape = 4, scale =.7), type = "l")
plot(seq(0, 20, .01), dgamma(seq(0, 20, .01)-5, shape = 4, rate =1.5), type = "l")


plot(seq(0, 20, .01), dbeta(seq(0, 20, .01)-5, 1.5, 1.5), type = "l")

```


```{r}

for(i in 1:nrow(priors)){
  prior_data[, priors$var[i]] <- (dbeta(prior_data$x, 2, 2) * (priors$max_prior[i] - priors$min_prior[i])) + priors$min_prior[i]
}
for(i in grep("Q.", names(prior_data), value = T)){
  eval(parse(text = paste0("prior_plot <- prior_plot + geom_line(aes(y = prior_data[,'", i, "']), group = '", i, "')")))
}

prior_plot_Q

plot(seq(0, 20, .01), dbeta(seq(0, 20, .01)-5, 1.5, 1.5), type = "l")


 # (("&$A3&"hyper * "&($G3 - $F3)&") + "&$F3&");"



priors <- readxl::read_xlsx("../03_estimate_weights/AUDIT_item_intervals.xlsx",
                            n_max = 18) %>% 
  transmute(var, label, min_prior = as.numeric(min_prior), 
            max_prior = as.numeric(max_prior))

prior_data <- data.frame(list(x = seq(0, 20, .01))) 



plot(1, 1, xlim = c(0, 20), ylim = c(0, 1))
for(i in 7:13){
  # prior_data[, priors$var[i]] <- (dbeta(prior_data$x, 2, 2) * (priors$max_prior[i] - priors$min_prior[i])) + priors$min_prior[i]
  lines(seq(priors$min_prior[i]+0.01, priors$max_prior[i]-.01, length.out = 99), qbeta(seq(0.01, .99, length.out = 99), 2, 2)* (priors$max_prior[i] - priors$min_prior[i]) + priors$min_prior[i] )
}
```


