---
title: "Concurrent validity of an Estimator of Weekly Alcohol Consumption (EWAC) based on the Extended AUDIT"
output: html_notebook
bibliography: bibliography.bib
link-citations: true
---


```{r, include=F}
# Loading data
source("../0 functions/functions.R")
source("../0 functions/filepaths.R")
library(data.table)
library(haven)
library(scales)

load(file.path(atsfiles, "qfmodule.RData"))
ats <- merge(ats, audit1, "audit1_label", all.x = T, all.y = F)
ats <- merge(ats, audit2, "audit2_label", all.x = T, all.y = F)
ats <- merge(ats, audit3, "audit3_label", all.x = T, all.y = F)
ats <- proc.AUDIT(ats)
ats <- ats[audit1 %in% 2:6 & audit2 %in% 1:7,]


```


## Background

The Alcohol Use Disorders Identification Test (AUDIT) [@Babor2001] is a 10-item questionnaire schedule widely employed in clinical practice and clinical research as a diagnostic test for alcohol use disorder (F10.1/F10.2). Diagnostic studies show good predictive capabilities for both:

-	clinical diagnoses (e.g. Composite International Diagnostic Interview, Structured Clinical Interview, Alcohol Use Disorder and Associated Disabilities Interview), and
-	consumption in excess of maximum recommended intakes (e.g. 140g/week in Australia, 112 g/week in UK)

in a variety of clinical and community dwelling populations [@DeMeneses-Gaya2009]. A variant of the AUDIT known as the ‘Extended AUDIT’ improves the granularity of the information collected on alcohol consumption. It follows the same schedule as the AUDIT, but provides a greater range of response options in the quantity and frequency items (questions 1 and 2). The Extended AUDIT has been used in UK research as part of two trials [@Kaner2013c, @Crane2018]  and the Alcohol Toolkit Study [@Beard2015a] to capture greater information on the higher risk drinkers, based on the observation that consumption items are right-truncated.

Neither the AUDIT nor the Extended AUDIT provide a measure of usual alcohol consumption, which could help individuals understand, monitor and adjust their drinking behaviour. Existing research on quantity-frequency alcohol instruments suggests that the product of the quantity and frequency questions in the AUDIT (see AUDIT-1 and AUDIT-2 in [Table 1](#table1)) could be used to produce a measure of usual consumption [@Lemmens1992].

The aim of the present study is develop an algorithm producing an Estimator of usual Weekly Alcohol Consumption in units (EWAC) using Extended AUDIT data as an input.



## Methods 

Table 1:  <a name="table1"></a>Overview of alcohol schedule used in this paper.

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th>Survey module</th>
<th>Schedule</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td><em>ATS QF module</em></td>
<td><ul>
<li>On how many days, if any, did you personally drink a drink containing alcohol in the last four weeks?</li>
<li>What was the maximum number of units you personally consumed on any one day when drinking an alcoholic drink or drinks in the last four weeks?</li>
<li>On how many days, if any, in the last four weeks did you personally drink… [prompting in turn &#39;51-60 units?&#39;, &#39;41-50 units?&#39;, …, &#39;1-2 units?</li>
</ul>
</tr>
<tr class="even">
<td align="center"><em>Health Survey for England</em></td>
<td><ul>
<li>Thinking now about all kinds of drinks, how often have you had an alcoholic drink of any kind during the last 12 months? <em>[8 items from &#39;Almost every day&#39; to &#39;Not at all in the last 12 months&#39;]</em></li>
<li>Did you have an alcoholic drink in the seven days ending yesterday?</li>
<li>On how many days out of the last seven did you have an alcoholic drink?</li>
<li>Which day last week did you last have an alcoholic drink/have the most to drink?</li>
<li>Thinking about last [answer to previous question], what types of drink did you have that day? [list of 8 types of alcohol beverages]</li>
<li>[running through each type of beverage and recording number of units drunk]</li>
</ul></td>

</tr>
</tbody>
</table>



## Results

### Sample sizes

```{r}
plot(ats$qfv,  ats$gfmeanweekly, pch =16, col = alpha(1, .2),
     ylab = "GF (units/week)", xlab = "EWAC (units/week)", main = NULL)
abline(0,1, col = 2, lwd =1.5)
```

