---
title: Concurrent validity of an Estimator of Weekly Alcohol Consumption (EWAC) based
  on the Extended AUDIT
csl: elsevier-vancouver.csl
link-citations: yes
always_allow_html: yes
output:
  html_document:
    df_print: paged
    number_sections: yes
    toc: yes
  pdf_document: 
    citation_package: natbib
    df_print: kable
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    toc: yes
bibliography: bibliography.bib
---

<!-- 
abstract	250
intro	456
methods	
results	
discussion	858
conclu	100
total	1664

-->

<!-- # Abstract -->
<!-- *Background and Aims:* -->
<!-- *Design:* -->
<!-- *Setting:* -->
<!-- *Participants/Cases:* -->
<!-- *Measurements:* -->
<!-- *Findings:* -->
<!-- *Conclusions:* -->


```{r loading data, include=F}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
# Loading data
source("../01_src/functions.R")
source("../01_src/filepaths.R")
source("../01_src/audit_weights.R")


knit_output_pdf <- F
if(knit_output_pdf){
  options(knitr.table.format = "pdf")
} else {
  options(knitr.table.format = "pandoc")
}


library(tidyverse)
library(knitr)
library(captioner)
library(kableExtra)
# library("scales")
# library("xtable")
# library("stargazer")

auditx_labels <- c("audit1_label", "audit2_label", "audit3_label")

load(file.path("..", "02_data", "HSE.rda"))
load(file.path("..", "02_data", "ATS.rda"))
# Too small a sample size too manage
ats$sex[ats$sex == "In another way"] <- NA
ats <- ats %>% 
  mutate(., 
    audit1_label = if_else(.$audit1_label %in% c("Refused", "Don't know"), 
                           NA_character_, audit1_label),
    audit2_label = if_else(.$audit2_label %in% c("Refused", "Don't know"), 
                           NA_character_, audit2_label),
    audit3_label = if_else(.$audit3_label %in% c("Refused", "Don't know"), 
                           NA_character_, audit3_label))

# Alcohol retail data from PHE 2017
alcohol_sales_2014_liters <- 361429387
alcohol_pcapita_adults_2014_units <-  16.3 # per capita for all population
alcohol_pcapita_drinker_2014_units <- 19.3 # per capita for drinkers only
MYPE_2014 <- 42724917 # population aged 18+



table_nums <- captioner::captioner(prefix = "Table")
tab_comparison_schedules <- table_nums(name = "tab_comparison_schedules",
                        caption = "Alcohol schedules: selective comparison")
tab_overview_datasets <- table_nums(name = "tab_overview_datasets",
                        caption = "Description of data sources")
tab_ewac_validity_reg <- table_nums(name = "tab_ewac_validity_reg",
                   caption = "Coefficients of linear regression of the bias and error of EWAC compared with GF")

# Hospital data

load("../02_data/hospital.rda")
vname_audit_baseline <- c("audit1_bline", "audit2_bline", "audit3_bline")
vname_audit_fu <- c("audit1_fu", "audit2_fu", "audit3_fu")
qdata <- records$main 

# Compute AUDIT-C scores
qdata$auditc_score <- proc_AUDIT_score(
  qdata[,vname_audit_baseline], audit_weights_version3)
qdata$auditc_risk <- proc_AUDIT_risk(qdata$auditc_score)

# Compute EWAC
qdata$ewac_mp <- proc_EWAC(qdata[,vname_audit_baseline],
                           audit_weights_midpoints, method = "qfv")
qdata$ewac_qfv <- proc_EWAC(qdata[,vname_audit_baseline],
                            audit_weights_STAN, method = "qfv" )
tdata_b <- records$tlfb %>%
  filter(encounter == "Baseline" &
           pid %in% unique(qdata$pid))
tdata_b <- merge(tdata_b, qdata, by = "pid")

```


# Introduction

Alcohol consumption is a significant cause of death and ill health. A number of strategies for prevention [@NICE-PH24], treatment and harm reduction [@NICE-CG115]. Many of those rely on the use of a screening tool. In the UK, the reference tool is the Alcohol Use Disorders Identification Test (AUDIT) [@Babor2001], a 10-item questionnaire schedule widely employed in clinical practice and clinical research as a diagnostic test for alcohol use disorder (ICD-10 F10.1/F10.2; ICD-11 QE10/6C40.1). The short, 3-item consumption version known as AUDIT-C has good diagnostics accuracy for both: (a) interview-based clinical diagnoses; and (b) consumption in excess of maximum recommended intakes (e.g. 140g/week in Australia, 112 g/week in UK) in a variety of populations [@DeMeneses-Gaya2009].

A variant of the AUDIT-C schedule known as the ‘Extended AUDIT-C’ improves the granularity of the information collected, thanks to a greater range of response options on quantity and frequency (items 1 and 2). The Extended AUDIT-C has been used in UK research as part of two trials [@Kaner2013c, @Crane2018] and one continuous household survey [@Beard2015a] to capture greater information on the higher risk drinkers, based on the observation that AUDIT consumption items are right-truncated.

Neither the AUDIT-C nor the Extended AUDIT-C provide a measure of usual alcohol consumption, which could help individuals understand, monitor and adjust their drinking behaviour. Making individuals more aware of the consumption and associated risk is a key recommendation in the recent prevention literature [@Nutt2014, @Rehm2016] and governmental health recommendations to the public [@AU.alcoguidelines2009, @AlcoholCMO2016b].

The aim of the present study is develop and validate an Estimator of usual Weekly Alcohol Consumption in units (EWAC) based on the Extended AUDIT-C. The product of frequency of drinking (AUDIT-C item 1) and quantity of drinking (AUDIT-C item 2) can be used to estimate usual alcohol consumption. In addition, measures of heavy use (AUDIT-C item 3) have been found to improve the consistency of the estimator with other data sources in previous research [@Lemmens1992].

This paper reports three distinct empirical studies. Study 1 estimates coefficients to apply to each of the AUDIT-C response items to compute an EWAC using data from a large English household survey, the Alcohol Toolkit Study (ATS). It then tests the concurrent validity of the EWAC across a number of demographic subpopulations in England. Study 2 tests the contingent validity of the EWAC compared with the 28-day Timeline Followback in a clinical population (visitors of an acute hospital). Study 3 compares the population-wide total and empirical cumulative distribution of alcohol consumption in England using the EWAC, the ATS graduated frequency schedule, the Health Survey for England, and official statistics on alcohol sales. The article concludes on the EWAC's potential applications in research and clinical care.

# Methods

## Data sources
A longstanding obstacle in alcohol research and care is the absence of a diagnostic gold standard, which is not addressed even by the development of new biomarkers. Instead, a number of instruments exist which measure self-reported alcohol consumption with varying validity and reliability. A state-of-the art review from @Greenfield2000 is summarised in [Table 1](#table1) below.

`r table_nums("tab_comparison_schedules")`

```{r table_comparison_schedules, fig.cap = tab_comparison_schedules, message=F}
readr::read_csv("table_overview_schedules.csv")  %>% 
  kable(., format = ifelse(knit_output_pdf, "latex", "html")) %>% 
  kable_styling()
```


Prospective diaries tend to record higher alcohol consumption by minimising recall bias, followed by GF, while lower levels seem to be recorded with QF measures [@Heeb2005, @Rehm1998]. 

The present paper employs four sources of data. An overview is available in [Table 2](#table2) below, together with references to publicly available methodological descriptions. Analysis of HSE and ATS microdata was approved by the University of Southampton's Faculty of Medicine Ethics Committee (ERGO 44682). The hospital questionnaire (Study 3) collection and analysis was approved by the Health Research Authority National Research Ethics Service (IRAS 247458;  REC 18/SC/0564). All results are reported in UK alcohol units (8g or 10mL of pure alcohol). Analyses are conducted in `R` [@RCoreTeam2017] using packages `tidyverse`, `survey`, `rstan` [@package-tidyverse; @package-rstan; @package-survey; @Lumley2004]. Computer programmes for all analyses are available on an online repository.

`r table_nums("tab_overview_datasets")`

|	Source	|	Waves	|	Population of inference	|	Alcohol schedules	|	Validation	|	Sample size	|	Design	|
|	-------------	|	-------------	|	-------------	|	-------------	|	-------------	|	------:	|	-------------	|
|	Alcohol Toolkit Study (ATS) [@Beard2015a]	|	November 2015-October 2017 (waves 110--133)	|	Residents of private English households aged 16 years and over	|	Extended AUDIT-C and GF	|	-	|	40,832	|	TBC	|
|	Health Survey for England (HSE) [@NatCenSocialResearch2013]	|	2011	|	Residents of private English households aged 18 years and over	|	Beverage-specific QF	|	This UK-specific schedule is consistent with a prospective diary [@Boniface2014] and yesterday recall [@Stockwell2016].	|	8,610	|	Multistage random sampling	|
|	Hospital questionnaire study (HOSP)	|	February 2019 to ??? 2019	|	Outpatient and day case hospital patients aged 18 years and over	|	Extended AUDIT-C and TLFB	|	TLFB records fewer drinking days and thus lower consumption than a prospective diary [@Grant1995]. This recall bias increases with the [@Hoeppner2010; @Vinson2003]	|	TBC	|	Cross-sectional recruitment with block randomisation to (a) self-administered AUDIT-C or (b) researcher-administered AUDIT-C. More detail in section 2.2.	|
|	Alcohol retail sales [@PHE2017]	|	2014	|	English population aged 18 years and over	|	N/A	|	See [@PHE2017].	|	N/A	|	Ratio of all alcohol produced or processed in the UK, as well as alcohol imported into the UK for sale and consumption, over the mid-year population estimate	|


## Study 1: EWAC estimation and concurrent validity

A pre-registered protocol for this analysis is available from @Dutey2018. We apply methods developed for quantity-frequency-variability instruments [@Lemmens1992] to the Extended AUDIT-C. For every individual $i$, we adjusting the product of $F_i$ and $Q_i$ (AUDIT items 1 and 2 respectively) with the frequency of intense drinking $V_i$ (AUDIT item 3):

$$\text{EWAC}_i = F_i Q_i + V_ib$$

where $b$ denotes the average number of units of alcohol consumed in an intense drinking day. Coefficients $F, Q, V$ and $b$ are unknown. In this study, two sets of potential coefficients are evaluated:

- using the AUDIT response item interval midpoint: for example, 2.5 for '2 to 3 times per week' in AUDIT item 2
- using a statistical model to estimate the coefficients directly from the ATS (further detail in supplementary material).

The ATS data is affected by missing data: `r tbrounding(mean(ats$audit1 == 1)*100)`% of respondents (`r formatbm(sum(ats$audit1 == 1))`) reported never drinking alcohol in AUDIT item 1 and were not asks any further AUDIT or GF questions.^[Note: Data from HSE on 'never-never?' indicates that a third are probably occasional drinkers. Consider adding the diagram previously drawn for ABH] They were thus excluded from the analysis. A further `r formatbm(sum(ats$audit1 != 1 & is.na(ats$gfmeanweekly)))` respondents (`r tbrounding(sum(ats$audit1 != 1 & is.na(ats$gfmeanweekly))/sum(ats$audit1 != 1))`% of those reporting drinking in AUDIT item 1) did not have a valid GF alcohol consumption record and were also excluded. Missing data in the ATS' GF schedule is assumed to be missing at random conditionally on the Extended AUDIT-C responses.

The discrepancy between the EWAC and the GF schedule is quantified, by treating GF as a gold standard. It enables the examination of two types of deviations across every participant $i$ in the ATS sample:

- MD: the mean deviation $\sum_{i=1}^{n}{n^{-1}(\rm{EWAC_i} - \rm{GF_i} )}$ can be assimilated to a measure of bias under the assumption that GF is a gold standard.
- RMSD: the root mean squared deviation $\sum_{i}{\sqrt{( \rm{EWAC_i} - \rm{GF_i} )^2}}$ can be assimilated to a measure of total error: it capture both bias and random deviation from the assumed gold standard.

We test the hypothesis that the EWAC's validity varies across population subgroups by regressing both (a) the simple deviation and (b) the root squared deviation in linear regression models, against the following predictors: sex by age group; ethnic group; favourite drink (wine, beer, etc.); highest educational qualification; religion; smoking status; general health; recent attempts at reducing or stopping drinking.

## Study 2: Concurrent validity in hospital outpatients

Participants were recruited from a range of clinics at a large acute hospital in Southampton, UK: orthopaedics outpatient, endoscopy day cases, young adult outpatient, infusion ???. 

<!-- discuss problems raised by Horton2007 -->


## Study 3: Aggregate concurrent validity

We examine the aggregate consistency in alcohol consumption estimates across all residents of England aged 18 years and over by plotting the empirical cumulative distribution of alcohol consumption given by (1) the EWAC estimated from the ATS; (2) the quantity-frequency estimated in the ATS; (3) the beverage-specific estimators in HSE in 2011; (4) the prospective diary estimator in HSE 2011. In this analysis, survey weights are used: in (1-3), poststratification weights estimated using calibration and age-sex MYPES; in (4), similar postratification weights adjusted for self-selection into participation to the prospective diary data collection. Finally, we examine the percentage of total alcohol sales for England accounted for by each method. Alcohol sales (both on-trade and off-trade) estimates for England (population aged 18 years and over) in 2014 were obtained from Public Health England [@PHE2017].


# Results

```{r}

ats$ewac_mp <- proc_EWAC(ats[, auditx_labels],
                           audit_coef = audit_weights_midpoints, method = "qfv")
ats$ewac_qf <- proc_EWAC(ats[, auditx_labels],
                           audit_coef = audit_weights_STAN, method = "qf")
ats$ewac_qfv <- proc_EWAC(ats[, auditx_labels],
                           audit_coef = audit_weights_STAN, method = "qfv")
ats$ewac_qb <- proc_EWAC(ats[, auditx_labels],
                           audit_coef = audit_weights_STAN, method = "qb")
ats$ewac_qv <- proc_EWAC(ats[, auditx_labels],
                           audit_coef = audit_weights_STAN, method = "qv")
ats$ewac_qfv_M <- proc_EWAC(ats[, auditx_labels],
                           audit_coef = audit_weights_STAN_men, method = "qfv")
ats$ewac_qfv_F <- proc_EWAC(ats[, auditx_labels],
                           audit_coef = audit_weights_STAN_women, method = "qfv")
ats <- mutate(ats, ewac_qfv_S = if_else(sex == "Men", 
                                        ewac_qfv_M, ewac_qfv_F))

ats$ewac_best <- data.frame(list(ats$gfmeanweekly - ats$ewac_qf, ats$gfmeanweekly - ats$ewac_qv, ats$gfmeanweekly - ats$ewac_qfv, ats$gfmeanweekly - ats$ewac_qb)) %>% 
  apply(., 1, function(X) if(length(which.min(abs(X)))==0) NA else which.min(abs(X))) 

ats$ewac_best<- factor(ats$ewac_best, labels = c("qf", "qv", "qfv", "qb"))

# table(ats$ewac_best, ats$ewac_qf <= ats$ewac_qfv)
test_midpoint <- list(
  MD = mean(ats$ewac_mp - ats$gfmeanweekly, na.rm = T),
  MD.p = t.test(ats$ewac_mp - ats$gfmeanweekly)$p.value,
  RMSD = sqrt(mean((ats$ewac_mp - ats$gfmeanweekly)^2, na.rm = T)),
  RMSD.p = t.test( (ats$ewac_mp - ats$gfmeanweekly)^2, mu = 4)$p.value
)

test_stan <- list(
  MD = mean(ats$ewac_qfv - ats$gfmeanweekly, na.rm = T),
  MD.p = t.test(ats$ewac_qfv - ats$gfmeanweekly)$p.value,
  RMSD = sqrt(mean((ats$ewac_qfv - ats$gfmeanweekly)^2, na.rm = T)),
  RMSD.p = chisq.test.variance( (ats$ewac_qfv - ats$gfmeanweekly)^2,
                                sigma = 4)$p.value
)

test_stan_S <- list(
  MD = mean(ats$ewac_qfv_S - ats$gfmeanweekly, na.rm = T),
  MD.p = t.test(ats$ewac_qfv_S - ats$gfmeanweekly)$p.value,
  RMSD = sqrt(mean((ats$ewac_qfv_S - ats$gfmeanweekly)^2, na.rm = T)),
  RMSD.p = chisq.test.variance( (ats$ewac_qfv_S - ats$gfmeanweekly)^2,
                                sigma = 4)$p.value
)

stan_qfv_S_men <- ats %>% 
  filter(sex == "Men") %>% 
  summarise(MD = mean(ewac_qfv_S - gfmeanweekly, na.rm = T),
            MD.p = t.test(ewac_qfv_S - gfmeanweekly)$p.value,
            RMSD = sqrt(mean((ewac_qfv_S - gfmeanweekly)^2, na.rm = T)),
            RMSD.p = chisq.test.variance((ewac_qfv_S - gfmeanweekly)^2,
                                         sigma = 4)$p.value) 
stan_qfv_S_women <- ats %>% 
  filter(sex == "Women") %>% 
  summarise(MD = mean(ewac_qfv_S - gfmeanweekly, na.rm = T),
            MD.p = t.test(ewac_qfv_S - gfmeanweekly)$p.value,
            RMSD = sqrt(mean((ewac_qfv_S - gfmeanweekly)^2, na.rm = T)),
            RMSD.p = chisq.test.variance((ewac_qfv_S - gfmeanweekly)^2,
                                         sigma = 4)$p.value)

```
## Study 1: EWAC estimation and validation

### EWAC coefficients

The first step involved choosing a set of coefficients to compute the EWAC. We compare using the midpoint of the AUDIT item intervals; and coefficients estimated using a statistical model. 
The EWAC based on midpoint coeffients produces a mean deviation (MD) of `r tbrounding(test_midpoint$MD, 2)` alcohol units (`r pvrounding(test_midpoint$MD.p)`), suggesting a slight positive bias. The root mean squared deviation (RMSD) is estimated at `r tbrounding(test_midpoint$RMSD, 2)`, significantly grater than 4 (`r pvrounding(test_midpoint$RMSD.p)`), suggesting that the EWAC of any given individual would be on average 12 units from the truth (treating GF schedule can be used as a gold-standard). 

Coefficients estimated by the statistical model provide a small improvement, but bias remains statistically significant with MD = `r tbrounding(test_stan$MD, 2)` (`r pvrounding(test_stan$MD.p)`). Error remains statistically significantly greater than 4 with  RMSD = `r tbrounding(test_stan$RMSD, 2)` (`r pvrounding(test_stan$RMSD.p)`). [Figure 1](#figure1) compares individual EWAC and GF values. We note the significant departure of lines of best fit, whether estimated using ordinary least squares or using c-splines.


<!-- With early signs that validity was strongly dependent on gender, we estimated coefficients independently for each gender to test for the use of gender-specific coefficients. This analysis was not planned in the pre-registered protocol but appeared to be justified by differences in alcohol consumption between women and men. Using gender-specific coefficients, we found HIGHER BIAS... -->



```{r warning=F}
dplyr::select(ats, gfmeanweekly, ewac_qfv) %>% 
  na.omit %>% 
  ggplot(mapping = aes(x = gfmeanweekly, y = ewac_qfv)) +
  geom_point(alpha = .2) +
  geom_abline(aes(intercept = 0, slope = 1, col = "Identity", linetype = "Identity")) +
  geom_smooth(aes(col="OLS", linetype = "OLS"), method ='lm', formula=y~x) +
  geom_smooth(aes(col="Spline", linetype = "Spline"), 
              method = 'gam', formula = y~s(x, bs="cs")) +
  scale_colour_manual(name="Line of fit",
    values = c("Identity" = "black", "OLS" = "blue", "Spline" = "red")) +
  scale_linetype_manual(name="Line of fit",
    values = c("Identity" = "solid", "OLS" = "longdash", "Spline" = "dotted")) +
  ylab("EWAC (UK units/week)") +
  xlab("Graduated-Frequency schedule (UK units/week)") +
  coord_equal() +
  scale_x_continuous(breaks = seq(0, 150, 10), limits = c(0, 150), minor_breaks = F)+
  scale_y_continuous(breaks = seq(0, 150, 10), limits = c(0, 150), minor_breaks = F)+
  geom_hline(yintercept = 14, color = "grey50") + 
  geom_vline(xintercept = 14, color = "grey50") +
  theme_bw() 

```

Figure 1: <a name="figure1"></a> Plot of EWAC against GF in ATS with ordinary least square regression line ($n$ = 22,373)


### Consistency across demographics


`r table_nums('tab_ewac_validity_reg')`

```{r, results='asis'}
library(stargazer)
lm_bias_mp <- lm((ewac_mp-gfmeanweekly) ~ sex*ageg + ethgrp + favdrink + 
                   tryalclyc2 + religion + highqual  + smokstat, data = ats)
lm_rmse_mp <- lm(sqrt((ewac_mp-gfmeanweekly)^2) ~ sex*ageg + ethgrp + favdrink + 
                   tryalclyc2 + religion + highqual  + smokstat, data = ats)
lm_bias_stan <- lm((ewac_qfv-gfmeanweekly) ~ sex*ageg + ethgrp + favdrink + 
                   tryalclyc2 + religion + highqual  + smokstat, data = ats)
lm_rmse_stan <- lm(sqrt((ewac_qfv-gfmeanweekly)^2) ~ sex*ageg + ethgrp + favdrink + 
                   tryalclyc2 + religion + highqual  + smokstat, data = ats)

stargazer(lm_bias_mp, lm_rmse_mp, 
          lm_bias_stan, lm_rmse_stan, 
          type = ifelse(knit_output_pdf, "latex", "html"), style = "ajs",
          ci = rep(T, 4), digits = 2, single.row = T, intercept.bottom = F, align=TRUE)

```


## Study 2: Concurrent validity in hospital patients

A total of `r nrow(qdata)` participants were recruited, with data successfully collected on the Extended AUDIT-C for `r round(mean(!is.na(qdata$auditc_score))*100)`% ($n$ = `r sum(!is.na(qdata$auditc_score))` of participants.


```{r}
library(lme4)
require(pscl)
tdata_b$qfv_log <- log(tdata_b$ewac_qfv/7)
tdata_b$mp_log <- log(tdata_b$ewac_mp/7)

mhosp <- glm(round(alcunits) ~ qfv_log + 
                 week_since, family = poisson, data = tdata_b)
mzip <- zeroinfl(round(alcunits) ~ qfv_log + tlfb_weekday +
                 week_since | week_since + tlfb_weekday, data = tdata_b)

summary(mhosp)


```


```{r}
scatterdata <- tdata_b %>% 
  filter(days_since <=28) %>% 
  group_by(pid) %>% 
  summarise(
    tlfb_units = mean(alcunits, na.rm = T)*7,
    ewac_mp = mean(ewac_mp),
    ewac_qfv = mean(ewac_qfv)) %>% 
  na.omit

ggplot(scatterdata, aes(y = ewac_qfv, x = tlfb_units)) +
  geom_point() +
  geom_abline() +
  geom_smooth(method='lm',formula=y~x) +
  theme_bw() +
  ylab("EWAC (UK units/week)") +
  xlab("Crude TLFB sample mean (UK units/week)")+
  coord_equal() +
  scale_x_continuous(breaks = seq(0, 100, 2), limits = c(0, 50), minor_breaks = F) +
  scale_y_continuous(breaks = seq(0, 100, 2), limits = c(0, 50), minor_breaks = F) +
  geom_hline(yintercept = 14, color = "grey50") + 
  geom_vline(xintercept = 14, color = "grey50")

```

## Study 3: Empirical distribution functions


```{r}
# From this point onwards, only drinkers aged 18 years and over

hse_dk18 <- filter(hse, age >= 18 & dnnow == 1) # only those drinking 'at all these days'
ats_dk18 <- filter(ats, actage >= 18 & audit1 > 1) %>%  # not 'never' drinkers
  mutate(agg_qfv = ewac_mp)
# ats_dk18$gfmeanweekly > 0 ???

```

*Table 3:* <a name="table3"></a>*Summary statistics on alcohol consumption in England (excluding abstainers)*

| Study     | Mean (units/week)     | Median | Variance | N | % of alcohol sold |
| ----------| ---------------------:| --------:|--------:| ----: | -----:
| HSE beverage-specific QF      | `r tbrounding(mean(hse_dk18$totalwu, na.rm =T))`   | `r tbrounding(median(hse_dk18$totalwu, na.rm =T))`     | `r tbrounding(var(hse_dk18$totalwu, na.rm =T))` |  `r formatbm(sum(!is.na(hse_dk18$totalwu)))` | `r tbrounding(mean(hse_dk18$totalwu, na.rm =T)*100/alcohol_pcapita_drinker_2014_units)` |
| HSE prospective diary | `r tbrounding(mean(hse_dk18$weektot, na.rm =T))`   | `r tbrounding(median(hse_dk18$weektot, na.rm =T))`     | `r tbrounding(var(hse_dk18$weektot, na.rm =T))` |  `r sum(!is.na(hse_dk18$weektot))` |`r tbrounding(mean(hse_dk18$weektot, na.rm =T)*100/alcohol_pcapita_drinker_2014_units)` |
| ATS GF    | `r tbrounding(mean(ats_dk18$gfmeanweekly, na.rm =T))`   | `r tbrounding(median(ats_dk18$gfmeanweekly, na.rm =T))`   | `r tbrounding(var(ats_dk18$gfmeanweekly, na.rm =T))` |  `r sum(!is.na(ats_dk18$gfmeanweekly))` |`r tbrounding(mean(ats_dk18$gfmeanweekly, na.rm =T)*100/alcohol_pcapita_drinker_2014_units)` |
| ATS EWAC    | `r tbrounding(mean(ats_dk18$agg_qfv, na.rm =T))`   | `r tbrounding(median(ats_dk18$agg_qfv, na.rm =T))`   | `r tbrounding(var(ats_dk18$agg_qfv, na.rm =T))` |  `r sum(!is.na(ats_dk18$agg_qfv))` |`r tbrounding(mean(ats_dk18$agg_qfv, na.rm =T)*100/alcohol_pcapita_drinker_2014_units)` |
| Retail sales | `r tbrounding(alcohol_pcapita_drinker_2014_units)` | -- | -- | -- | --|



```{r}
# ggplot() + 
#   stat_ecdf(aes(agg_qfv, colour = "ATS EWAC"), data = ats_dk18, geom = "step") +
#   stat_ecdf(aes(gfmeanweekly, colour = "ATS GF"), data = ats_dk18, geom = "step") +
#   stat_ecdf(aes(totalwu, colour = "HSE QF"), data = hse_dk18, geom = "step") +
#   stat_ecdf(aes(weektot, colour = "HSE Diary"), data = hse_dk18, geom = "step") +
#   ylab("Empirical cumulative density") +
#   xlab("Average units/week") +
#   scale_y_continuous(breaks = seq(0, 1, .1)) +
#   scale_x_continuous(limits = c(1, 100), 
#                      breaks = seq(0, 500, 10),
#                      minor_breaks = seq(0, 500, 5)) +
#   geom_vline(xintercept = 14)
#   
```

```{r echo=F, fig.height=5, fig.width=8}
library(latticeExtra)
library(scales)
xlims = c(-3, 153)
EWAC <- ecdfplot(na.omit(ats_dk18$agg_qfv),
                key= keyABH, panel = function(x, f.value = NULL, type = "s",
                                              groups = NULL, qtype = 7, ref = TRUE,...) {
                  panel.ecdfplot(x, f.value = NULL, type = "s",
                                 groups = NULL, qtype = 7,
                                 ref = TRUE,...)
                  panel.abline(v = 14)
                },
                xlim = xlims, col = alpha("orange", .7), lwd = 3, lty = 1,
                scales=list(cex=1.5), xlab = ecdfxlab, ylab = ecdfylab)

GF <- ecdfplot(na.omit(ats_dk18$gfmeanweekly),
                key= keyABH, panel = function(x, f.value = NULL, type = "s",
                                              groups = NULL, qtype = 7, ref = TRUE,...) {
                  panel.ecdfplot(x, f.value = NULL, type = "s",
                                 groups = NULL, qtype = 7,
                                 ref = TRUE,...)
                  panel.abline(v = 14)
                },
                xlim = xlims, col = alpha("black", .7), lwd = 3, lty = 1,
               scales=list(cex=1.5), xlab = ecdfxlab, ylab = ecdfylab)

HSE <- ecdfplot(na.omit(hse_dk18$totalwu),
                key= keyABH, panel = function(x, f.value = NULL, type = "s",
                                              groups = NULL, qtype = 7, ref = TRUE,...) {
                  panel.ecdfplot(x, f.value = NULL, type = "s",
                                 groups = NULL, qtype = 7,
                                 ref = TRUE,...)
                  panel.abline(v = 14)
                },
                xlim = xlims, col = alpha("red", .7), lwd = 1, lty = 1,
                scales=list(cex=1.5), xlab = ecdfxlab, ylab = ecdfylab)

HSEdiary <- ecdfplot(na.omit(hse_dk18$weektot),
                key= keyABH, panel = function(x, f.value = NULL, type = "s",
                                              groups = NULL, qtype = 7, ref = TRUE,...) {
                  panel.ecdfplot(x, f.value = NULL, type = "s",
                                 groups = NULL, qtype = 7,
                                 ref = TRUE,...)
                  panel.abline(v = 14)
                },
                xlim = xlims, col = alpha("red", .7), lwd = 2, lty = 2,
                scales=list(cex=1.5), xlab = ecdfxlab, ylab = ecdfylab)

HSEdiary + HSE +EWAC + GF

```


## Regression analysis

```{r echo =F}

#mbias <- lm(qfv-gfmeanweekly ~ as_factor(agez) + as_factor(sexz) + as_factor(favdrink), data = ats) #[gfmeanweekly <20,]
# ats$alcorisk <- cut(ats$auditc_score, breaks = c(1, 4, 8, 13))
# levels(ats$alcorisk) <- c("Low", "Increased", "Higher")
# mbias <- lm(qfv-gfmeanweekly ~ as_factor(ageg) + sex + favdrink + (alcorisk), data = ats)
#  tab3 <- stargazer(mbias, title="Bias regression results",
#                      align=TRUE, type = "html",style = "ajs")
#  
# merr <- lm((qfv-gfmeanweekly)^2 ~ as_factor(ageg) + sex + favdrink + (alcorisk), data = ats)
#  tab4 <- stargazer(mbias, title="MSE regression results",
#                      align=TRUE, type = "html",style = "ajs")
# #kable(tab2)
# kable(xtable(mbias))
#  kable(xtable(merr))
```



# Discussion

<!-- This paper adopted a QFV approach to compute the EWAC over arguably more direct approaches possible with the ATS, such as supervised machine learning or direct estimation of the mean alcohol consumption for every observed combination of Extended AUDIT-C responses. Although cumbersome, the deterministic nature of the QFV algorithm is meant to be easier to estimate with small training datasets. It is also easier to implement in relational databases employed in electronic record systems compared to many manchine learning techniques. -->


## Main findings

To the best of our knowledge, this is the first paper to develop an EWAC using a well-accepted and fast alcohol questionnaire such as the AUDIT. We used the Extended AUDIT-C to address the AUDIT-C's right-truncation and estimate consumption even for heavy consumers.

We find that it is possible to provide an EWAC with a margin of error of XX units using gender-specific coefficients, with consistent results across the English population.

Much like we produced gender-specific coefficients, it would be possible to use age information to adjust coefficients.
is consitent with previous research suggesting  largest discrepancy is noted for individuals aged 85 years and over

Knowledge of alcohol beverage content remains insufficient in the UK. In the 2007 HSE, just 31% of frequent beer drinkers and 17% of frequent wine drinkers were not aware of the unit equivalence to their intake. These proportions were found to be even lower in all other drinkers. 


This study reports an extensive range of validation analyses on the EWAC's validity and reliability. Findings are valid for the general population of England, although they may not generalise to other populations excluded from our sampling frames, such as residents of communal and carceral institutions, homeless people, or migrant populations. Similarly, findings may not designated to small subpopulations with an atypical alcohol consumption, such as patients seeking care for conditions such as addition or alcohol liver disease.


## Limitations 

As noted previously, a low proportion of the population knows the alcohol content of what they drink in units. This may influence the reliability of AUDIT item 2 (quantity), particularly in cases when it is self-administered or administered without guidance on beverage alcohol content. In this paper, the AUDIT scheduled was administered by a researcher in all cases but XXX participants in study 3; and even those were given a showcard summarising the unit content of most popular beverages. 
<!-- During data collection, we recorded a frequent confusion with the AUDIT's item 1: some participants being confused as to whether the question refers to number of days in which alcohol has been consumed, or number of occasions on any given day. We believe this to be a common occurrence when the AUDIT is self-administered. -->

respondents who drank beer and those who drank wine at least once a week were much more likely to know how many units were in that drink than were those who seldom drank these drinks, but even so, a of the



drinking diary was given to all participants aged 18 and over who completed the main HSE interview and had had an alcoholic drink in the previous 12 months.



<!-- The ATS presents some limitations. First, the high levels of data missingness and the exclusion of respondents answering 'Never' in AUDIT item 1 constituted a high risk of introducing bias in Study 1. Second, there is to our knowledge no validation data available for the ATS graduated-frequency schedule. For these reasons, basing the EWAC on ATS alone would have been hazardous. Such inferential risks are nevertheless addressed in the present paper through Study 3, through data collected from a health care setting, which low levels of missingness, using a different alcohol schedule known to exhibit more variability, but lower bias. -->

<!-- - GF not missing at random. Impute the `r round(3883/(3883+22376)*100, 1)`% missing cases -->



## Is the EWAC good enough to be used?

Our findings demonstrate that the mean deviation between EWAC and a more time-consuming alcohol schedule is XX (GF) and XX (TFLB). In other words, 95% of respondents' expected alcohol intake may be up to X units away from the EWAC. Other clinical parameters may be affected by similar uncertainty. For instance, blood pressure is highly variables and can be xx off using a single reading. 
Specific patient and public involvement is needed to understand how satisfactory this precision is. In the meantime, we argue that the EWAC can be useful to both clinicians and patients alongside established AUDIT-C thresholds: it can be used to engage a conversation around health risks of alcohol and recommendations to maintain consumption below 14 units in the UK. We also think it can be less stigmatising than a score in excess of 5, which can often be interpreted by patients as being labelled as alcohol dependent or problem drinker.

## What are the potential applications?

Important contributions by Nutt and Rehm invite us to reconsider alcohol care practice in favour of a more clinical approach. These authors argue that alcohol consumption should be treated like any other risk factor: blood pressure or cholesterol are frequently monitored by general physicians in a non stigmatising way and managed as well as possible. Similarly, Rehm and Nutt's view is that  individuals should know their consumption, and general practitioners should be able to engage with patients effectively in managing risks. Where lifestyle interventions are unsuccessful, pharmacological alternatives should be sought. All recommendations participate to a single goal: preventing the very emergence of alcohol harm, by which state treatment is often less successful. 

Supportive of their argument, we propose that the EWAC could be employed in routine clinical practice to feedback results from an AUDIT-C check and provide recommendations to all types of drinkers; encouraging low-risk drinkers to maintain this lifestyle, and informing those close to 14 units or exceeding that limit that monitoring and controling alcohol consumption can be easy and improve long-term health. The EWAC can realistically only be computed by software and we have made such a calculator available online to this end: soton.ac.uk. It may be particularly suitable to digital intervention where personalised feedback can be generated as a function of not just mean alcohol consumption, but also intensity of alcohol consumption. For instance, some individuals drinking very rarely do so in large quantities in any given session. The EWAC and the AUDIT item 3 can be used to provide more relevant advice on consuming no more than 5 units in any given session. Study 3 showed in clinical population that the EWAC seemed to be as reliable when the Extended AUDIT-C is self-administered as when it is administered by a researcher. 
 
# Conclusion


# Ethics




# Tables


Table 1:  <a name="table_questionnaires"></a>Overview of alcohol schedule used in this paper.

<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th>Survey module</th>
<th>Schedule</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td><em>ATS QF module</em></td>
<td><ul>
<li>On how many days, if any, did you personally drink a drink containing alcohol in the last four weeks?</li>
<li>What was the maximum number of units you personally consumed on any one day when drinking an alcoholic drink or drinks in the last four weeks?</li>
<li>On how many days, if any, in the last four weeks did you personally drink… [prompting in turn &#39;51-60 units?&#39;, &#39;41-50 units?&#39;, …, &#39;1-2 units?</li>
</ul>
</tr>
<tr class="even">
<td align="center"><em>Health Survey for England</em></td>
<td><ul>
<li>Thinking now about all kinds of drinks, how often have you had an alcoholic drink of any kind during the last 12 months? <em>[8 items from &#39;Almost every day&#39; to &#39;Not at all in the last 12 months&#39;]</em></li>
<li>Did you have an alcoholic drink in the seven days ending yesterday?</li>
<li>On how many days out of the last seven did you have an alcoholic drink?</li>
<li>Which day last week did you last have an alcoholic drink/have the most to drink?</li>
<li>Thinking about last [answer to previous question], what types of drink did you have that day? [list of 8 types of alcohol beverages]</li>
<li>[running through each type of beverage and recording number of units drunk]</li>
</ul></td>

</tr>
</tbody>
</table>


# References